Fast Sorting on Modern Computers

### 2005-08-31 docs/tech_report/intro.tex
fixed small errors. didnt really get rid of any TODOs
made final pass

Papers considering sorting performance typically analyse instruction count.
LaMarca and Ladner \cite{LaMarca96} break this trend by discussing the cache
implications of several algorithms, including sorts. This report 
investigates these claims and tests and analyses high performance sorts
based upon thier work.

LaMarca devotes a chapter of his thesis to sorting. Another chapter is spent
discussing implicit heaps, which are later used in heapsort. In
\cite{LaMarca99}, radixsort was added for comparison. Radixsort is added here
for the same purpose. For comparison, elementary sorts - insertion sort,
selection sort, bubblesort and shakersort - are also included here.

@@

Several important discovered results are reported here. The
branch prediction rates of insertion sort and selection sort, for example, and
the difference in performance between binary searches and sequential searches due
to branch misses. Several steps to remove instructions from standard algorithms
are devised, such as double tiling mergesort, removing a sentinel and bounds
check from heapsort.

@@

Chapter \ref{conclusions} presents conclusions, and lists contributions
described by this report.

>>>>

Papers considering sorting performance typically analyse instruction count.
LaMarca and Ladner \cite{LaMarca96} break this trend by discussing the cache
implications of several algorithms, including sorts. This report investigates
these claims and tests and analyses high performance sorts based upon their
work.

LaMarca devotes a chapter of his thesis to sorting. Another chapter is spent
discussing implicit heaps, which are later used in heapsort. In
\cite{LaMarca99}, radixsort was added for comparison. Radixsort and shellsort
are added here for the same purpose. For comparison, elementary sorts -
insertion sort, selection sort, bubblesort and shakersort - are also included.

@@

Several important discovered results are reported here. The branch prediction
rates of insertion sort and selection sort, for example, and the difference in
performance between binary searches and sequential searches due to branch
misses. Several steps to remove instructions from standard algorithms are
devised, such as double tiling and aligning mergesort, removing a sentinel and
bounds check from heapsort, and adding a copy-back step to radixsort.

@@

Chapter \ref{conclusions} presents conclusions, summarises results from previous
chapters, compares our results with LaMarcas, presents a list of cache and
branch prediction results and improvements, discusses the fastest sort in our
tests and lists contributions made by this research.
